---
title: "LC-MS of *U.saggitifolia* in negative ionization mode"
author: "Pablo Corella, Jefferson Pastuña"
date: "2023-06-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
usethis::git_vaccinate()
```

## Introduction

The present document aims to record the procedure given for the statistical analysis of secondary metabolites present in the different growth stages of *Urospatha saggitifolia*. For each step a brief explanation, the code and graphics obtained are included.

The workflow used is taken from the paper ["notame": Workflow for Non-Targeted LC–MS Metabolic Profiling](https://doi.org/10.3390/metabo10040135). Which offers a wide variety of functions to perform metabolomic profile analysis.

## Before to start

The "notame" package accepts as input a feature table that can be obtained through software such as MZMine, MSDial, among others.

In this case, the table was obtained with the help of both programs. The data was processed in MZmine to obtain the ideal analysis parameters as well as to confirm the presence of previously found positive controls (known metabolites). Then, the obtained parameters were entered in MS-Dial to generate a (*.txt) file that was slightly modified to obtain the feature table.

Modifications made to the raw (*.txt) file can be summarized in adding and renaming columns. The added columns "Column" and "Ion Mode" allow to analyze samples with different types of columns and with different ionization modes respectively. Also, the cells corresponding to mass and retention time must be renamed so that the package can detect and process it.

## Notame workflow

As a first step for the analysis, all the necessary libraries were installed and loaded in Rstudio.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}
library(notame)  
library(doParallel) 
library(magrittr)  
library(tidyverse) 
library(patchwork) 
library(dplyr)
```

Then, a log system was added to have a record of each process executed.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
ppath <- "C:/Users/F4ss0/Documents/Ikiam21062022/Tesis-Pablo-Corella/U_sagittifolia_tubers/"
init_log(log_file = paste0(ppath, 'Result/LC-MS_Negative/log_LC-MS_Negative.txt'))
```

Next, the feature list was imported using the read_from_excel function

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
NEG_data <- read_from_excel(file = paste0(ppath, "Data/NEG_feature_list.xlsx"), sheet = 1, corner_row = 4, corner_column = "D", split_by = c("Column", "Ion mode"))
```

Once the data was read, the next step was to create a MetaboSet in order to create a specific R object.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
NEG_modes <- construct_metabosets(exprs = NEG_data$exprs, pheno_data = NEG_data$pheno_data, feature_data = NEG_data$feature_data, group_col = "Group")
```

Finally, each mode was extracted in a single object.

```{r}
NEG_mode <- NEG_modes$RP_NEG
```

As a additional step, we can visualize the raw data in order to inspect the processing routines.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
NEG_raw_sambx <- plot_sample_boxplots(NEG_mode, order_by = "Group")
NEG_raw_pca <- plot_pca(NEG_mode, center = T)
NEG_raw_pca + NEG_raw_sambx
```

In the boxplot we can see how the abundance of metabolites present in the QCs does not vary between them, so the drift correction could be optional.

## Preprocessing

The first step of the preprocessing is to change the features with value equal to 0 to NA.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
NEG_mode <- mark_nas(NEG_mode, value = 0)
```

Then, features with low detection rate are first flagged and then will be removed. The notame package employs two criteria to select this features. First, is the feature presence in a percentage of QC injections, and then the feature presence in a percentage within a sample group or class.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
NEG_mode <- flag_detection(NEG_mode, qc_limit = 0.75, group_limit = 0.8)
```

With these values, features which that were not detected in the 75% of the QC injections and 80% of sample groups will be discarded.

The next steps for preprocessing correspond to drift correction. The drift correction can be applied by smoothed cubic spline regression.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
NEG_dc <- dc_cubic_spline(NEG_mode)
NEG_corrected <- NEG_dc$object
NEG_corrected <- flag_quality(NEG_corrected)
```

Then we can visualize the correction for QCs.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
NEG_corr_sambx <- plot_sample_boxplots(NEG_corrected, order_by = "Group")
NEG_corr_pca <- plot_pca(NEG_corrected, center = T) 
NEG_corr_pca + NEG_corr_sambx
```

Finally, a step to consider in preprocessing is feature cleastering. This step helps us reduce the number of features of the same molecule that were split due to ionization problems or unexpected behavior.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}
NEG_clusteredQC <- cluster_features(NEG_corrected, rt_window = 1/60, all_features = T, corr_thresh = 0.9, d_thresh = 0.8)
NEG_compressedQC <- compress_clusters(NEG_clusteredQC)
```

Finally, we can plot the PCA.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
NEG_PCA_2<-plot_pca(NEG_compressedQC)
NEG_PCA_2
```

# Second PCA and loading plot

Droping flagged features

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
NEG_no_flag <- drop_flagged(NEG_compressedQC)

# Extracting feature table (Expression data)
NEG_feature_table <- exprs(NEG_no_flag)

# Extracting Phenotipic data
NEG_pheno_data <- NEG_no_flag@phenoData@data
```

Preparing data and transposing feature table.

```{r}
NEG_feat_table_pca <- t(NEG_feature_table)

#Changing NA to 0 
NEG_feat_table_pca[is.na(NEG_feat_table_pca)]=0

# Centering and Scaling features
NEG_pca_noQC <- prcomp(NEG_feat_table_pca, center = T, scale. = T)
```

Plotting PCA results.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
NEG_scores <- NEG_pca_noQC$x %>%         # Get PC coordinates
  data.frame %>%                         # Convert to data frames
  mutate(Sample_ID = rownames(.)) %>%    # Create a new column with the sample names
  left_join(NEG_pheno_data)              # Adding metadata
  
ggplot(NEG_scores, aes(PC1, PC2, shape = Group, color = Group)) +
  geom_point(size = 3) +
  guides(x=guide_axis(title = "PCA 1 (40%)"), y=guide_axis(title = "PCA 2 (23%)")) +
  theme_classic()
```  

Plotting loading results.

```{r}
NEG_loadings <- NEG_pca_noQC$rotation %>%    # Extract loadings
  data.frame(Feature_name = rownames(.))     # New column with feat name
```

Creating an artificial table with Feature name and Compound column.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
NEG_feat_name <- readxl::read_excel("NEG_Iden_Comp_4.xlsx", 1)

# Creating a new small table of the annotated compounds
NEG_compouds_all <- left_join(NEG_feat_name, NEG_loadings)

# Plotting results
ggplot(NEG_loadings, aes(PC1, PC2)) + 
  geom_point(alpha = 0.1) +
  theme_classic() + 
  geom_point(data = NEG_compouds_all, aes(shape = Identification_level, color = Identification_level), size = 3) +
  ggrepel::geom_label_repel(data = NEG_compouds_all, aes(label = Compound), max.overlaps = 40,
                               box.padding = 0.3) +
  guides(x=guide_axis(title = "PC 1 (37%)"), y=guide_axis(title = "PC 2 (24%)")) +
  ggsci::scale_color_aaas()
```

# Heat map plot

Loaded of identified metabolites abundance from Excel file.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}
met_hetmap <- readxl::read_excel("NEG_Data_Pablo_to_Hetmap.xlsx", 3)
met_hetmap <- met_hetmap[order(met_hetmap$A1),]     # sorted by column
row.names(met_hetmap) <- met_hetmap$X               # Named rows
met_hetmap_m <- data.matrix(met_hetmap)             # DataFrame
met_hetmap_m1 <- met_hetmap_m[1:35,-1]              # Delete the old rows with name
```

Logarithmic scale.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}
log_met_hetmap_m1 <- log10(met_hetmap_m1)
y <- log_met_hetmap_m1
```

Plotting Heat map

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}
# Package installation and library loading
#if (!requireNamespace("devtools", quietly = TRUE)) {
#  install.packages("devtools")
#}
#devtools::install_github("cran/gplots")
library(gplots)

#Clusters
hr <- hclust(as.dist(1 - cor(t(y), method = "pearson")), method = "complete")
hc <- hclust(as.dist(1 - cor(y, method = "spearman")), method = "complete")
mycol <-  colorpanel(75, "blue", "white", "red")

#Creación del mapa de calor
heatmap.2( y, Rowv = as.dendrogram(hr), Colv = as.dendrogram(hc), col = mycol, density.info = "none", trace = "none", dendrogram = "both", scale = "row", labRow = NULL, labCol = NULL, margins = c(4, 14), key = TRUE, keysize = 1.5, key.xlab = NA, key.ylab = NA)
```

Save Heatmap plot.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}
png(filename ="LC_MS2_Neg_HeatMap.png", width = 14, height = 8.64, units = "in", pointsize = 12, res=300)
heatmap.2(y, Rowv = as.dendrogram(hr), Colv = as.dendrogram(hc), col = mycol, density.info = "none", trace = "none", dendrogram = "both", scale = "row", labRow = NULL, labCol = NULL, margins = c(4, 14), key = TRUE, keysize = 0.5, key.xlab = NA, key.ylab = NA)
dev.off() 
```

Finish a record.

```{r}
finish_log()
```

