---
title: "Statistic of total metabolite content in *U. sagitifolia*"
author: "Jefferson Pastuna"
date: "2023-12-20"
output:
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
usethis::git_vaccinate()

```

# Introduction

The present document aims to record the procedure given for the statistical analysis of secondary metabolites quantification in the different growth stages of *Urospatha saggitifolia*. For each step a brief explanation, the code and graphics obtained are included.

## Data loading

Installation of R packages and loading of the libraries.

```{r library, echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}

# Loadding ggplot2 library
library(ggplot2)

# agricolae package installation and library loadding
#install.packages("agricolae", repos = "https://cran.r-project.org")
library(agricolae)

# ggbreak agricolae package installation and library loadding
remotes::install_github("YuLab-SMU/ggbreak")
library(ggbreak)

```

Loaded Excel data.

```{r data}

# Data of total phenolic content
total_phenolic <- data.frame(readxl::read_excel("../Data/Total_Metabolite_Content.xlsx", 1))

# Data of total flavonoid content
total_flavonoid <- data.frame(readxl::read_excel("../Data/Total_Metabolite_Content.xlsx", 2))

```

## Boxplot

Before performing the ANOVA test we can inspect the data using boxplot. Next we extract the data for each growth stage.

```{r factor}

# Total phenolic content in different growth stage
s_phe <- total_phenolic[c(1:3),1]
j_phe <- total_phenolic[c(1:3),2]
a_phe <- total_phenolic[c(1:3),3]

# Total flavonoid content in different growth stage
s_fla <- total_flavonoid[c(1:3),1]
j_fla <- total_flavonoid[c(1:3),2]
a_fla <- total_flavonoid[c(1:3),3]

```

Later we grouped the data to plotting boxplot.

```{r samples}

# Sample length
n_sample  <- length(s_phe)

# To total phenolic content
cont_phe <- c(s_phe,j_phe,a_phe)
rep_phe <- c(rep("Seedling",n_sample),rep("Juvenile",n_sample),rep("Adult",n_sample))
datos_phe <- data.frame(rep_phe,cont_phe)

# To total flavonoid content
cont_flav <- c(s_fla,j_fla,a_fla)
rep_flav <- c(rep("Seedling",n_sample),rep("Juvenile",n_sample),rep("Adult",n_sample))
datos_flav <- data.frame(rep_flav,cont_flav)

```

Plotting boxplot.

```{r box}

# To total phenolic content
bx_phe <- ggplot(datos_phe,aes(x=rep_phe,y=cont_phe,fill=rep_phe)) +
  geom_boxplot() +
  ggtitle("Boxplot of total phenolic content") +
  guides(x=guide_axis(title = NULL),
         y=guide_axis(title = "Total phenolic content \n(mg Gallic acid/g Extract)"),
         fill=guide_legend(title="Growth stage")) +
  theme(axis.text.x=element_text(angle=90, hjust=1))

# To total flavonoid content
bx_flav <- ggplot(datos_flav,aes(x=rep_flav,y=cont_flav,fill=rep_flav)) +
  geom_boxplot() +
  ggtitle("Boxplot of total flavinoid content") +
  guides(x=guide_axis(title = NULL),
         y=guide_axis(title = "Total flavonoid content \n(mg Quercetin/g Extract)"),
         fill=guide_legend(title="Growth stage")) +
  theme(axis.text.x=element_text(angle=90, hjust=1))

bx_phe
bx_flav

```

## Bartlett test

Previus ANOVA test.

```{r bartlett1}

# To total phenolic content
bartlett.test(cont_phe~rep_phe,data = datos_phe)

```

Because of p-value is 0.5407, and its mayor to 0.05, we accept (H0). Therefore, samples have same varinces.

```{r bartlett2}

# To total flavonoid content
bartlett.test(cont_flav~rep_flav,data = datos_flav)

```

Because of p-value is 0.2058, and its mayor to 0.05, we accept (H0). Therefore, samples have same varinces.

## ANOVA test

ANOVA model.

```{r logit}

# Logit model to total phenolic content
mylogit_phe <- glm(cont_phe~rep_phe,data = datos_phe)

# Logit model to total flavonoid content
mylogit_flav <- glm(cont_flav~rep_flav,data = datos_flav)

```

Print ANOVA test.

```{r anova}

# To total phenolic content
anova(mylogit_phe,test="LRT")

# To total flavonoid content
anova(mylogit_flav,test="LRT")

```

## Tukey test

```{r tukey}

# To total phenolic content
modelo_phe <- aov(cont_phe~rep_phe,data = datos_phe)
tukey_phe <- HSD.test(modelo_phe,"rep_phe", group=TRUE,console=TRUE)
plot(tukey_phe, cex.names=0.7, ylab = "Media diferences", main="Tukey to total phenolic content", font.main=3)

# To total flavonoid content
modelo_flav <- aov(cont_flav~rep_flav,data = datos_flav)
tukey_flav <- HSD.test(modelo_flav,"rep_flav", group=TRUE,console=TRUE)
plot(tukey_flav, cex.names=0.7, ylab = "Media diferences", main="Tukey to total flavonoid content", font.main=3)

```

